\mytitle

### 2015-08-06 MSc_diss.tex
Almost done

\begin{abstract}
Abstract.300 words.
\end{abstract}

\tableofcontents
\newpage

@@

A study that enabled us to re-examine the theory of $O_3$ and $RONO_2$ formation processes in a quite unique environment has been conducted by \cite{Neuman2012}. After the Deepwater Horizon (DWH) offshore oil platform accident on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface. Subsequent oil evaporation caused a release of methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment. At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas). Daytime airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$\footnote{Peroxyacetyl nitrate.}+$RONO_2$) and VOCs in the marine boundary layer downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated ozone formation chemistry in the plume. This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6--10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating ozone production. As a result, on 10 June, enhancements ozone levels in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.

@@

The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on VOCs concentration (therefore called $VOC$-sensitive) (e.g., \cite{Fowler2008}). When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere. Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle. In other words, the following reaction sequence predominates:
\begin{equation}
O_3 + hv \rightarrow O(^1D) + O_2\tag{\ref{reac:O3+hv=O1D+O2}}
\end{equation}
\begin{equation}
O(^1D) + H_2O \rightarrow OH + OH\tag{\ref{reac:O1D+H2O=OH+OH}}
\end{equation}
\begin{equation}
O(^1D) \xrightarrow{M} O(^3P)\tag{\ref{reac:O1D=O3P}}
\end{equation}
\begin{equation}
O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
\end{equation}
\begin{equation}
OH + CO \rightarrow HO_2 + CO_2\tag{\ref{reac:OH+CO=H+CO2}}
\end{equation}
\begin{equation}\label{reac:OH+RH=RO2+H2O}
OH + RH \rightarrow RO_2 + H_2O
\end{equation}
\begin{equation}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2\tag{\ref{reac:HO2+HO2=H2O2+O2}}
\end{equation}
\begin{equation}\label{reac:RO2+HO2=ROOH+O2}
RO_2 + HO_2 \rightarrow ROOH + O_2
\end{equation}
\begin{equation}
O_3 + HO_2 \rightarrow OH + 2O_2\tag{\ref{reac:HO2+O3=OH+2O2}}
\end{equation}
At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}). $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high VOCs concentrations respectively). $NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}). $CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}). Mixing ratios of other species in our model are of the order of magnitude as in ACLBM.

@@

When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:RO2+NO=RO+NO2}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals. The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO=H+CO2}) and (\ref{reac:OH+RH=RO2+H2O})), therefore making this regime $VOC$-sensitive. Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
\begin{equation}
OH + NO_2 \xrightarrow{M} HNO_3\tag{\ref{reac:OH+NO2=HNO3}}
\end{equation}
Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases. However, if the concentrations of VOCs are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO=H+CO2}), (\ref{reac:OH+RH=RO2+H2O})) can compete with reaction (\ref{reac:OH+NO2=HNO3}), leading to an increase in the $O_3$ production rate.

Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of $VOC$-sensitive regime by our model is not expected. It also makes it difficult to split $NO_x$-sensitive and $VOC$-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions. However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels, for example, 5 ppb (Figure \ref{fig:model_validation}).

@@

The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and $VOC$ mentioned above is to construct a ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}. The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and $VOC$ concentrations. When $VOC$/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase. The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and $VOC$-sensitive regimes. Under the 'ridge line' the $VOC$/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to $VOC$. Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
	\centering
    \begin{subfigure}[t]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
        \caption{(ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies. The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h. Adapted from \citep{Sillman1999}.} \label{fig:Sillman1999}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.eps}
        \caption{as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e. the only $VOC$ is $CO$).} \label{fig:noAN_netO3rate}
    \end{subfigure}
    \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color)}\label{fig:netO3rate}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Rewrite!

The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off. Taking into account the fact that the maximum $NO_x$ mixing ratio considered in the experiment is 10 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed square. Moreover, since the minimum $NO_x$ mixing ratio considered in the experiment is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate is seen in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999}.

One can see the same pattern in these two figures. At high $VOC$/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and $VOC$ levels (which corresponds to the top right corner of the green dashed square). The change in $O_3$ production rate at high $VOC$/$NO_x$ ratios in our model is largely insensitive to $VOC$, which fits into the theory. At low $VOC$/$NO_x$ ratios the $O_3$ production rate starts to depend on $VOC$ concentration also resembling the behaviour of ozone isopleths at the top left corner of the green dashed square. The overall tendency of $O_3$ production rate to increase with increasing $VOC$ and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory. However, the modelled values of $O_3$ production rates are much lower than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 9 ppb/h, while in theory it is 30 ppb/h, meaning that our model underestimates the $O_3$ production rate, but reproduces some features of the relationship between $O_3$, $NO_x$ and $VOC$.

@@

To evaluate the sensitivity of $O_3$ production to alkyl nitrate chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out. In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.

As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between two reactions:
\begin{equation}
RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2\tag{\ref{reac:RO2+NO=aphaRO+NO2}}
\end{equation}
\begin{equation}
RO_2 + NO \xrightarrow{\alpha_3} RONO_2\tag{\ref{reac:RO2+NO=aphaRONO2}}
\end{equation}
The reaction (\ref{reac:RO2+NO=aphaRO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=aphaRONO2}) terminates this process. Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h] %  AN branching ratios
\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
\centering
\begin{tabular}{cc}
\hline
Alkyl nitrate    & Branching ratio \\
\hline
methyl           & 0.001 \\
ethyl            & 0.009 \\
iso-propyl       & 0.042 \\
n-propyl         & 0.020 \\
n-butyl          & 0.033 \\
sec-butyl        & 0.090 \\
iso-butyl        & 0.033 \\ 
tert-butyl       & 0.025 \\
n-pentyl         & 0.052 \\
2-pentyl         & 0.129 \\
3-pentyl         & 0.131 \\
2-methylbutyl    & 0.052 \\
3-methyl-2-butyl & 0.141 \\
2-methyl-2-butyl & 0.047 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] % netO3rate noAN withAN diff
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}%
    \hfill
\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle). Difference between them is shown on the right.}\label{fig:netO3rate_noAN_withAN_diff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h] % netO3mixrat noAN allAN diff diffpercent
\centering
\begin{minipage}{.45\textwidth} % noAN
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % diff
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % allAN
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % diffpercent
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb, range 250 ppt - 100 ppb) with alkyl nitrate chemistry switched off (a) and switched on (b). Absolute difference (with alkyl nitrates minus without them) between two series of model runs in ppb (c), same in \% (d). Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The contribution of different alkyl nitrates to the $O_3$ decrease varies: the largest drop --- 11.6 ppb --- is caused by ethyl nitrate ($C_2H_5ONO_2$), and is almost the same for methyl nitrate. On the other hand, iso-pentyl nitrates account for the smallest $O_3$ decrease.

>>>>

\begin{abstract}
This study attempts to investigate the role of alkyl nitrates in tropospheric ozone chemistry. A conceptual box model representing the chemistry of mean conditions of global troposphere was developed. The sensitivity of ozone production rate to inclusion of different alkyl nitrates into the set of reactions was analysed in various regimes of ozone precursors. For the first time the ozone production efficiency and its relation to alkyl nitrate presence was estimated using the so-called carbon bonds method. The model behaviour stays in accordance with previous research, while the major findings of the present study offer an insight into alkyl nitrate distribution in the $NO_x$-VOC space.
\end{abstract}

\newpage
\tableofcontents
\newpage

@@

A study that enabled us to re-examine the theory of $O_3$ and $RONO_2$ formation processes in a quite unique environment has been conducted by \cite{Neuman2012}. After the Deepwater Horizon (DWH) offshore oil platform accident on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface. Subsequent oil evaporation caused a release of methane and other light alkanes to the atmosphere forming a VOC mixture atypical for the remote marine environment. At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas). Daytime airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$\footnote{Peroxyacetyl nitrate.}+$RONO_2$) and VOCs in the marine boundary layer downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated ozone formation chemistry in the plume. This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6--10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating ozone production. As a result, on 10 June, enhancements ozone levels in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.

@@

The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on VOCs concentration (therefore called VOC-sensitive) (e.g., \cite{Fowler2008}). When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere. Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle. In other words, the following reaction sequence predominates:
\begin{equation}
O_3 + hv \rightarrow O(^1D) + O_2\tag{\ref{reac:O3+hv=O1D+O2}}
\end{equation}
\begin{equation}
O(^1D) + H_2O \rightarrow OH + OH\tag{\ref{reac:O1D+H2O=OH+OH}}
\end{equation}
\begin{equation}
O(^1D) \xrightarrow{M} O(^3P)\tag{\ref{reac:O1D=O3P}}
\end{equation}
\begin{equation}
O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
\end{equation}
\begin{equation}
OH + CO \rightarrow HO_2 + CO_2\tag{\ref{reac:OH+CO=H+CO2}}
\end{equation}
\begin{equation}\label{reac:OH+RH=RO2+H2O}
OH + RH \rightarrow RO_2 + H_2O
\end{equation}
\begin{equation}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2\tag{\ref{reac:HO2+HO2=H2O2+O2}}
\end{equation}
\begin{equation}\label{reac:RO2+HO2=ROOH+O2}
RO_2 + HO_2 \rightarrow ROOH + O_2
\end{equation}
\begin{equation}
O_3 + HO_2 \rightarrow OH + 2O_2\tag{\ref{reac:HO2+O3=OH+2O2}}
\end{equation}
At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}). $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high VOCs concentrations respectively). $NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}). $CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}). Mixing ratios of other species in our model are of the order of magnitude as in ACLBM.

@@

When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:RO2+NO=RO+NO2}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals. The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO=H+CO2}) and (\ref{reac:OH+RH=RO2+H2O})), therefore making this regime VOC-sensitive. Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
\begin{equation}
OH + NO_2 \xrightarrow{M} HNO_3\tag{\ref{reac:OH+NO2=HNO3}}
\end{equation}
Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases. However, if the concentrations of VOCs are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO=H+CO2}), (\ref{reac:OH+RH=RO2+H2O})) can compete with reaction (\ref{reac:OH+NO2=HNO3}), leading to an increase in the $O_3$ production rate.

Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of VOC-sensitive regime by our model is not expected. It also makes it difficult to split $NO_x$-sensitive and VOC-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions. However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels, for example, 5 ppb (Figure \ref{fig:model_validation}).

@@

The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and VOC mentioned above is to construct an ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}. The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and VOC concentrations. When VOC/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase. The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and VOC-sensitive regimes. Under the 'ridge line' the VOC/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to VOC. Above the 'ridge line' the VOC/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing VOC and decreases with increasing $NO_x$ \citep{Sillman2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
	\centering
    \begin{subfigure}[t]{0.45\linewidth}
        \centering
        \includegraphics[width=\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
        \caption{for mean summer daytime meteorology and clear skies. The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h. Adapted from \citep{Sillman1999}.} \label{fig:Sillman1999}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.eps}
        \caption{from the first series of experiments with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e. the only VOC is $CO$).} \label{fig:noAN_netO3rate}
    \end{subfigure}
    \caption{Isopleths of net rate of $O_3$ production (ppb/h) as a function of VOC (ppbC) and $NO_x$ (ppb).}\label{fig:netO3rate}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first series of experiments with alkyl nitrate chemistry switched off. Taking into account the fact that the minimum sum of VOCs mixing ratios considered in this work is 1782 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed rectangle. Moreover, since the minimum $NO_x$ mixing ratio considered in this work is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate occurs in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999} (i.e., not all of our modelling data is shown in Figure \ref{fig:noAN_netO3rate}).

One can see the same pattern in these two figures. At high VOC/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and VOC levels. In theory at high $NO_x$ and VOC levels the $O_3$ production rate decreases (top right corner of green rectangle), due to impact of $HNO_3$ chemistry. Therefore, this figure confirms once again that the model does not account for impact of $HNO_3$ chemistry.

The change in $O_3$ production rate at high VOC/$NO_x$ ratios in our model is somewhat insensitive to VOC, which fits into the theory. At low VOC/$NO_x$ ratios the $O_3$ production rate starts to depend on VOC concentration also resembling the behaviour of ozone isopleths at the middle of green rectangle. The overall tendency of $O_3$ production rate to increase with increasing VOC and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory. The modelled values of $O_3$ production rates are similar than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 25 ppb/h, in theory it is 30 ppb/h, meaning that our model shows the same $O_3$ production rates, but does not show the same distribution of $O_3$ production rate in $NO_x$-VOC space as in the theory.

@@

To evaluate the sensitivity of $O_3$ production to alkyl nitrate chemistry under different $NO_x$ and VOC conditions a number of simulations were carried out. In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.

As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between two reactions:
\begin{equation}
RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2\tag{\ref{reac:RO2+NO=aphaRO+NO2}}
\end{equation}
\begin{equation}
RO_2 + NO \xrightarrow{\alpha_3} RONO_2\tag{\ref{reac:RO2+NO=aphaRONO2}}
\end{equation}
The reaction (\ref{reac:RO2+NO=aphaRO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=aphaRONO2}) terminates this process. Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h!] %  AN branching ratios
\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
\centering
\begin{tabular}{cc}
\hline
Alkyl nitrate    & Branching ratio \\
\hline
methyl           & 0.001 \\
ethyl            & 0.009 \\
iso-propyl       & 0.042 \\
n-propyl         & 0.020 \\
n-butyl          & 0.033 \\
sec-butyl        & 0.090 \\
iso-butyl        & 0.033 \\ 
tert-butyl       & 0.025 \\
n-pentyl         & 0.052 \\
2-pentyl         & 0.129 \\
3-pentyl         & 0.131 \\
2-methylbutyl    & 0.052 \\
3-methyl-2-butyl & 0.141 \\
2-methyl-2-butyl & 0.047 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] % netO3rate noAN withAN diff
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.eps}
        \caption{}
    \end{subfigure}%
    \hfill
\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of VOC (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (top left) and switched on (top right). Difference between them is shown at the bottom.}\label{fig:netO3rate_noAN_withAN_diff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h] % netO3mixrat noAN allAN diff diffpercent
\centering
\begin{minipage}{.45\textwidth} % noAN
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % diff
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % allAN
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
\end{minipage}
\begin{minipage}{.45\textwidth} % diffpercent
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb_new.eps}
\end{minipage}
\caption{Isopleths of net $O_3$ production (ppb, color) as a function of VOC (ppbC) and $NO_x$ (ppb, range 250 ppt - 100 ppb) with alkyl nitrate chemistry switched off (top left) and switched on (top right). Absolute difference between two series of model runs in ppb (bottom left), same in \% (bottom right). Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The biggest change in $O_3$ (11.6 ppb) is related to ethyl nitrate, though its production by the end of the simulation is the smallest in comparison to other nitrates. Furthermore, when these is no conversion of methane to methyl nitrate, but conversion of ethane to ethyl nitrate is present even at a small initial value of ethane ($C_2H_6$ 3.8 ppb in comparison to 1915 ppb of methane) we observe almost the same change in $O_3$ mixing ratio. Surprisingly, in experiment with full alkyl nitrate chemistry we see almost the same $O_3$ change as in case of methyl or ethyl nitrate.
